An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2016-09-29 nips_2016.tex
almost done

Our approach is to instead decompose $X_{\rm log}$ as
\begin{equation}\label{eq:deconvolution}
    X_{\rm log} = X_{\rm norm}+X_{\rm corr},
\end{equation}
where we assume $X_{\rm norm} \sim \mathcal{N}(0, \sigma^2)$ and that $X_{\rm
corr}$ is a zero-mean ``correction'' variable with density $C_{\sigma}(X)$.  The
two variables are added (i.e., their distributions convolve) to form $X_{\rm
log}$.  This decomposition requires an appropriate $X_{\rm corr}$ distribution.
We show in Section~\ref{sec:correction} that it is possible to use deconvolution
to derive a highly accurate representation of $C_\sigma(X)$.
Figure~\ref{fig:deconvolution} shows three examples of $X_{\rm norm}$ and
$X_{\rm corr}$ densities that sum to $X_{\rm log}$.

Using $X_{\rm corr}$ samples from $C_\sigma(X)$, the acceptance test is now
\begin{equation}\label{eq:criteria}
    \Delta + X_{\rm log} = (\Delta + X_{\rm norm}) + X_{\rm corr} = \Delta^* + X_{\rm corr} >0.
\end{equation}
Therefore, assuming the variance of $\Delta^*$ is small enough, if we have an
estimate of $\Delta^*$ from the current data minibatch, we test acceptance by
adding a random variable $X_{\rm corr}$ and then accept $\theta'$ if the result
is positive (and reject otherwise).

@@

Our proposed test in Equation~\ref{eq:criteria} requires knowing the
distribution of the correction variable $X_{\rm corr}$ such that $X_{\rm norm} +
X_{\rm corr} = X_{\rm log}$, where $X_{\rm norm} \sim \mathcal{N}(0,\sigma^2)$
and $X_{\rm log}$ has a standard logistic CDF, $(1+\exp(-X))^{-1}$. In
Section~\ref{sec:analysis}, we show that the accuracy of the test depends on the
absolute error between the CDFs of $X_{\rm norm} + X_{\rm corr}$ and $X_{\rm
log}$. Consequently, we need to minimize this in our construction of $X_{\rm
corr}$. More formally, let $\Phi_\sigma(X) = \Phi(X/\sigma)$ where $\Phi$ is the
standard normal CDF\footnote{Hence, $\Phi_\sigma$ is the CDF of a zero-mean
Gaussian with variance $\sigma$.}, $S(X)$ be the logistic function, and
$C_{\sigma}(X)$ be the \emph{density} of the correction $X_{\rm corr}$
distribution. Our goal, based on Equation~\ref{eq:deconvolution}, is to solve
the following optimization problem:
\begin{equation}\label{eq:overall_corr_problem}
    C_\sigma^* = \argmin_{C_\sigma} |\Phi_{\sigma} * C_{\sigma} - S|
\end{equation}
where $*$ denotes convolution.
% {\color{blue} Daniel: This is very unusual use of
%convolutions (i.e., CDF * DENSITY = CDF). It's valid but I was very surprised to
% know that.}
For computation of $C_\sigma$, we assume that its input $Y$ and
another variable $X$ lie in the intervals $[-V,V]$ and $[-2V,2V]$, respectively.
% {\color{blue} Daniel: I am a little confused at the different interval sizes.}
In continuous form, the optimization problem is therefore
\begin{equation}\label{eq:conv1}
    C_\sigma^* = \argmin_{C\sigma} \; \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|
\end{equation}
To get $C_\sigma^*$ in a computable form, we discretize $X$ and $Y$ into $4N+1$
and $2N+1$ values respectively. If $i \in \{-2N, \ldots, 2N\}$ and $j \in \{-N,
\ldots, N\}$, then we can write $X_i = i(V/N)$ and $Y_j = j(V/N)$, and the
condition in Equation~\ref{eq:conv1} can be re-expressed as:
\begin{equation}\label{eq:conv2}
    C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \{-2N,\ldots,2N\}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.
\end{equation}
To make the problem easier to understand, we can define a matrix $M$ and vectors
$u$ and $v$ such that $M_{ij} = \Phi_{\sigma}(X_i-Y_j)$, $u_j =
C_{\sigma}(Y_j)$, and $v_i = S(X_i)$, where the indices $i$ and $j$ are
appropriately translated to be non-negative indices for $M, u,$ and $v$. Thus,
Equation~\ref{eq:conv2} is equivalent to minimizing the $L_{\infty}$ norm:
\begin{equation}\label{eq:optimization_l1}
    u^* = \argmin_{u} \|Mu-v\|_{\infty}.
\end{equation}
% {\color{blue} Daniel: Equation~\ref{eq:optimization_l1} is finding
% the $C_\sigma$ minimizing the $L_{\infty}$ norm, but the previous one,
% Equation~\ref{eq:conv2}, is finding the $C_\sigma$ that is minimizing the
% $L_\infty$ norm, i.e., the maximum value! If we want this to be the $L_{\infty}$ norm
% (and I think we do) then Equation~\ref{eq:conv2} needs an outer sum over the $i$
% terms. In addition, Equation~\ref{eq:conv1} also needs to be modified, perhaps
% eliminate the ``sup'' but keep $x \in [-2V,2V]$}

@@

\SetKwInOut{Input}{Input}
\SetKwInOut{Output}{Output}
\begin{algorithm}[t]
\Input{Number of samples $T$, minibatch size $m$, error bound $\delta$, pre-computed correction $C_1(X)$ distribution, initial sample $\theta_1$.}
\Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta_T\}$ from $p(\theta)$\;}
\For{$t=\{1, \ldots, T\}$}{
    Propose a candidate $\theta'$\ from proposal distribution $q(\theta'\mid\theta_t)$\;
    Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(\theta_t,\theta')$ and sample variance $\sigma^2(\Delta^*)$\;
    Estimate moments $E|\Lambda_i-\Lambda|$ and $E|\Lambda_i-\Lambda|^3$ from the sample, and error $\epsilon$ from Equation~\ref{eq:clt-bounds3}\;
    \While {$\sigma^2(\Delta^*) \ge 1$ {\bf or} $\epsilon > \delta$}{
    	Draw $m$ more samples to augment the minibatch, update $\Delta^*$, $\sigma^2(\Delta^*)$ and $\epsilon$ estimates\;
    }
    Draw $X_{\rm norm} \sim \mathcal{N}(0,1-\sigma^2(\Delta^*))$ and
    $X_{\rm corr}$ from the correction distribution $C_1(X)$\;

        \eIf{$\Delta^* + X_{\rm norm} + X_{\rm corr} > 0$}{
            Accept the candidate, $\theta_{t+1} = \theta'$\;
        }{
            Reject and re-use the old sample, $\theta_{t+1} = \theta_t$\;
        }

@@

{\color{blue} Daniel: TODO improve/describe Table~\ref{tab:xcorr} and try to
relocate Algorithm~\ref{alg:our_algorithm} to be somewhere nicer.}

@@

For the analysis, in Section~\ref{ssec:delta_star_distribution}, we first
present bounds on the absolute and relative error (in terms of the CDFs) of the
distribution of $\Delta^*$ vs a Gaussian. We then show in
Section~\ref{ssec:preserve_bounds} that these error bounds are preserved after
the addition of other random variables, in particular $X_{\rm norm}$ and $X_{\rm
corr}$. From this it follows that the acceptance test has the same error bound.

@@

{\color{blue} Daniel: the following lemma is, I think, from Corollary 2
in~\cite{explicit-clt05} but I cannot derive the exact lemma as we have it
here.} To start, we use the following quantitative central-limit result
(from~\cite{explicit-clt05}):

\begin{lemma}\label{lem:quant_clt}
Let $X_1,\ldots,X_n$ be a set of zero-mean, independent, identically-distributed
random variables with sample mean $\hat{X}$ and variance $\sigma^2(\hat{X})$
where:
\begin{equation}
    \hat{X} = \frac{1}{n}\sum_{i=1}^nX_i \qquad {\rm and} \qquad \sigma(\hat{X}) = \frac{1}{n}\left(\sum_{i=1}^n(X_i-\hat{X})^2\right)^{\frac{1}{2}}.
\end{equation}
This means $t=\hat{X}/\sigma(\hat{X})$ has an approximate Student's distribution
which approaches a normal distribution in the limit. Then
from~\cite{explicit-clt05}:
\begin{equation}\label{eq:clt-bounds}
    \sup_x|{\rm Pr}(t<x) - \Phi(x)| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}}.
\end{equation}
\end{lemma}

{\color{blue} Daniel: the notation here was (and still is, to a lesser extent)
the source of considerable confusion for me. Do we want $\sigma(X)$ or
$\sigma(\hat{X})$? I think in Lemma~\ref{lem:quant_clt} we want the former, not
the latter. OH ALSO, this section uses $n$ as the full data size, which means
before this section, we should use $n$, not $N$, for the full data size, right?
The Bardenet paper uses $n$ to represent the full data size.

PS: I'd recommend not putting sample mean and sample variance in an equation if
the $\sigma(\hat{X})$ gets clarified, since those computations are common
knowledge.

PPS: The sample mean is probably more commonly denoted as $\bar{X}$.}

Lemma~\ref{lem:quant_clt} demonstrates that as long as we know the first and
third absolute moments $E|X|$ and $E|X|^3$, we can bound the error of the normal
approximation, which decays as $O(n^{-\frac{1}{2}})$. Making the change of
variables $y = x \sigma(\hat{X})$, Equation~\ref{eq:clt-bounds} becomes
\begin{equation}\label{eq:clt-bounds2}
   \sup_y\left|{\rm Pr}(\hat{X}< y) - \Phi\left(\frac{y}{\sigma(\hat{X})}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}},
\end{equation}
showing that the distribution of $\hat{X}$ approaches the normal distribution
$\mathcal{N}(0,\sigma^2(\hat{X}))$ whose variance is $\sigma^2(\hat{X})$,
measured from the sample.

To apply this to our test, let $X_i = \Lambda_i(\theta,\theta') -
\Lambda(\theta,\theta')$, so that the $X_i$ are zero-mean, i.i.d. variables. If
instead of all $n$ samples, we only extract a subset of $b$ samples
corresponding to our minibatch, we can connect $\hat{X}$ with our $\Delta^*$
term:
\begin{equation}\label{eq:x-hat}
    \hat{X} = \Delta^*(\theta,\theta') - \Delta(\theta,\theta'),
\end{equation}
so that $\sigma(\hat{X}) = \sigma(\Delta^*)$. {\color{blue} Daniel: This time it
makes sense to use $\sigma(\hat{X})$ instead of $\sigma(X)$.} This results in
the following Corollary:

\begin{corollary}\label{cor:our_bound_delta_prime}
We can now substitute into Equation~\ref{eq:clt-bounds2} and displace by the mean, giving:
\begin{equation}\label{eq:clt-bounds3}
    \sup_y\left|{\rm Pr}(\Delta^*< y) - \Phi\left(\frac{y-\Delta}{\sigma(\Delta^*)}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{b}} = \epsilon(\theta,\theta',b)
\end{equation}
\end{corollary}

Corollary~\ref{cor:our_bound_delta_prime} shows that the distribution of
$\Delta^*$ approximates a Normal distribution with mean $\Delta$ and variance
$\sigma^2(\Delta^*)$. Furthermore, it bounds the error with \emph{estimable
quantities}: both $E|X|$ and $E|X|^3$ can be determined as means of $|\Lambda_i
- \Lambda|$ and $|\Lambda_i - \Lambda|^3$, respectively, on each minibatch. We
expect this will often be accurate enough on minibatches with hundreds or
thousands of points, but otherwise bootstrap CIs can be computed from those
sequences. Since the bounds are monotone in $E|X|$ and $E|X|^3$, using upper
bootstrap CI limits will provide high-confidence error bounds.

@@

{\color{blue} Daniel: Hold on, the following Corollary seems to add $X_{\rm norm}$
to both random variables we are comparing (and \emph{then} $X_{\rm corr}$). But
didn't the previous section already justify that $\Delta^* \approx \Delta +
X_{\rm norm}$? That is, I don't think we need to add another $X_{\rm norm}$; I
think we just want to add $X_{\rm corr}$. Otherwise, this new test does not
match what we wrote earlier.

I think the easiest way to resolve this is to do the following: we assume that
$\Delta^*$ is already $\Delta + X_{\rm noise}$. Then all we need to do is add
$X_{\rm corr}$. So this is the same as what we are doing, except we do not need
to apply Lemma~\ref{lem:cdf_bounds} twice; we just need to do it once because we
are only adding $X_{\rm corr}$, not $X_{\rm norm}$.}

@@

\begin{corollary}\label{cor:bounds_preserved}
If $\sup_y|{\rm Pr}(\Delta^* < y) - \Phi(\frac{y-\Delta}{\sigma(\Delta^*)})|
\leq \epsilon(\theta,\theta',b)$, then
\begin{equation}
    \sup_y|{\rm Pr}(\Delta^*+X_{\rm norm}+X_{\rm corr} < y) - S(y-\Delta)| \leq \epsilon(\theta,\theta',b)
\end{equation}
where $S(x)$ is the standard logistic function, and $X_{\rm norm}$ and $X_{\rm corr}$ are generated as per Algorithm 1. 
\end{corollary}

\begin{proof}
We apply Lemma~\ref{lem:cdf_bounds} twice. First take:
\begin{equation}
    P(y) = {\rm Pr}(\Delta^* < y) \qquad {\rm and}\qquad Q(y) = \Phi\left(\frac{y-\Delta}{\sigma(\Delta^*)}\right)
\end{equation}
and convolve with the distribution of $X_n$ {\color{blue} (Daniel: to be clear I
think this was supposed to be $X_{\rm norm}$ instead of $X_n$ but see comments
earlier about why I think it is not necessary.)} which has density
$\phi(X/\sigma_n)$ where $\sigma_n^2 = 1 - \sigma^2(\Delta^*)$. This yields the
next iteration of $P$ and $Q$:
\begin{equation}
    P'(y) = {\rm Pr}(\Delta^*+X_{\rm norm} < y) \qquad {\rm and}\qquad Q'(y) = \Phi\left({y-\Delta}\right)
\end{equation}
Now we convolve with the distribution of $X_{\rm corr}$ which gives:
\begin{equation}
    P''(y) = {\rm Pr}(\Delta^*+X_{\rm norm}+X_{\rm corr} < y) \qquad {\rm and}\qquad Q''(y) = S\left({y-\Delta}\right)
\end{equation}
Both steps preserve the error bound $\epsilon(\theta,\theta',b)$. Finally
$S(y-\Delta)$ is a logistic CDF centered at $\Delta$, and so $S(y-\Delta) = {\rm
Pr}(\Delta + X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We conclude
that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm
norm}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm
log} > 0)$ by at most $\epsilon$.
\end{proof}

@@

{\color{blue} Daniel: Do we want to include this new analysis? There seem to be
two different ways we show the bound. The first part (above) is nice and I like
it (and it seems to show all we need for Algorithm 1, doesn't it?}

@@

Here the limit $G_n(x,y)$ is a normal CDF plus a correction term that decays as
$n^{1/2}$. {\color{blue} (Daniel: how did the equation in the next sentence
appear? I verified it and it seems correct, but it looks like it popped out of
nowhere.)} Importantly, since $x^2\phi(x) = \phi^{''}(x) + \phi(x)$ where
$\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
\begin{equation}\label{eq:GNderivatives}
    G_n(x,y) = \Phi(x) + \frac{y}{6\sqrt{n}}(2\phi^{''}(x)+3\phi(x))
\end{equation}
From which we see that $G_n(x,y)$ is a linear combination of $\Phi(x)$,
$\phi(x)$ and $\phi^{''}(x)$. This is quite fortuitous. In Algorithm 1, we
correct for the difference in $\sigma$ between $\Delta^*$ and the variance
needed by $X_{\rm corr}$ using $X_{\rm norm}$. This same method works when we
wish to estimate the error in $\Delta^*$ vs $G_n(x,y)$. Since all of the
component functions of $G_n(x,y)$ are derivatives of a (unit variance)
$\Phi(x)$, adding a normal variable with variance $\sigma'$ increases the
variance of all three functions to $1+\sigma'$. Thus we add $X_{\rm norm}$ as
per Algorithm 1 preserving the limit in Equation~\ref{eq:GNderivatives}.

>>>>

Our approach is to instead decompose $X_{\rm log}$ as
\begin{equation}\label{eq:deconvolution}
    X_{\rm log} = X_{\rm norm}+X_{\rm corr},
\end{equation}
where we assume $X_{\rm norm} \sim \mathcal{N}(0, \sigma^2)$ and that $X_{\rm
corr}$ is a zero-mean ``correction'' variable with density $C_{\sigma}(X)$.  The
two variables are added (i.e., their distributions convolve) to form $X_{\rm
log}$.  This decomposition requires an appropriate $X_{\rm corr}$ distribution.
We show in Section~\ref{sec:correction} that it is possible to use deconvolution
to derive a highly accurate representation of $C_{\sigma}(X)$.
Figure~\ref{fig:deconvolution} shows three examples of $X_{\rm norm}$ and
$X_{\rm corr}$ densities that sum to $X_{\rm log}$.

Using $X_{\rm corr}$ samples from $C_{\sigma}(X)$, the acceptance test is now
\begin{equation}\label{eq:criteria}
    \Delta + X_{\rm log} = (\Delta + X_{\rm norm}) + X_{\rm corr} = \Delta^* + X_{\rm corr} >0.
\end{equation}
Therefore, assuming the variance of $\Delta^*$ is small enough, if we have an
estimate of $\Delta^*$ from the current data minibatch, we test acceptance by
adding a random variable $X_{\rm corr}$ and then accept $\theta'$ if the result
is positive (and reject otherwise).

@@

Our proposed test in Equation~\ref{eq:criteria} requires knowing the
distribution of the correction variable $X_{\rm corr}$ such that $X_{\rm norm} +
X_{\rm corr} = X_{\rm log}$, where $X_{\rm norm} \sim \mathcal{N}(0,\sigma^2)$
and $X_{\rm log}$ has a standard logistic CDF, $(1+\exp(-X))^{-1}$. In
Section~\ref{sec:analysis}, we show that the accuracy of the test depends on the
absolute error between the CDFs of $X_{\rm norm} + X_{\rm corr}$ and $X_{\rm
log}$. Consequently, we need to minimize this in our construction of $X_{\rm
corr}$. More formally, let $\Phi_{s_X} = \Phi(X/s_X)$ where $\Phi$ is the
standard normal CDF\footnote{Hence, $\Phi_{s_X}$ is the CDF of a zero-mean
Gaussian with variance $s_X$.}, $S(X)$ be the logistic function, and
$C_{\sigma}(X)$ be the \emph{density} of the correction $X_{\rm corr}$
distribution. Our goal, based on Equation~\ref{eq:deconvolution}, is to solve
the following optimization problem:
\begin{equation}\label{eq:overall_corr_problem}
    C_\sigma^* = \argmin_{C_\sigma} |\Phi_{\sigma} * C_{\sigma} - S|
\end{equation}
where $*$ denotes convolution.
% {\color{blue} Daniel: This is very unusual use of
%convolutions (i.e., CDF * DENSITY = CDF). It's valid but I was very surprised to
% know that.}
For computation of $C_\sigma$, we assume that its input $Y$ and
another variable $X$ lie in the intervals $[-V,V]$ and $[-2V,2V]$, respectively.
% {\color{blue} Daniel: I am a little confused at the different interval sizes.}
In continuous form, the optimization problem is therefore
\begin{equation}\label{eq:conv1}
    C_\sigma^* = \argmin_{C\sigma} \; \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|
\end{equation}
To get $C_\sigma^*$ in a computable form, we discretize $X$ and $Y$ into $4N+1$
and $2N+1$ values respectively. If $i \in \{-2N, \ldots, 2N\}$ and $j \in \{-N,
\ldots, N\}$, then we can write $X_i = i(V/N)$ and $Y_j = j(V/N)$, and the
condition in Equation~\ref{eq:conv1} can be re-expressed as:
\begin{equation}\label{eq:conv2}
    C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \{-2N,\ldots,2N\}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.
\end{equation}
To make the problem easier to understand, we can define a matrix $M$ and vectors
$u$ and $v$ such that $M_{ij} = \Phi_{\sigma}(X_i-Y_j)$, $u_j =
C_{\sigma}(Y_j)$, and $v_i = S(X_i)$, where the indices $i$ and $j$ are
appropriately translated to be non-negative indices for $M, u,$ and $v$. Thus,
Equation~\ref{eq:conv2} is equivalent to minimizing the $L_{\infty}$ norm:
\begin{equation}\label{eq:optimization_l1}
    u^* = \argmin_{u} \|Mu-v\|_{\infty}.
\end{equation}
% {\color{blue} Daniel: Equation~\ref{eq:optimization_l1} is finding
% the $C_\sigma$ minimizing the $L_{\infty}$ norm, but the previous one,
% Equation~\ref{eq:conv2}, is finding the $C_\sigma$ that is minimizing the
% $L_\infty$ norm, i.e., the maximum value! If we want this to be the $L_{\infty}$ norm
% (and I think we do) then Equation~\ref{eq:conv2} needs an outer sum over the $i$
% terms. In addition, Equation~\ref{eq:conv1} also needs to be modified, perhaps
% eliminate the ``sup'' but keep $x \in [-2V,2V]$}

@@

\SetKwInOut{Input}{Input}
\SetKwInOut{Output}{Output}
\begin{algorithm}[t]
\Input{Number of samples $T$, minibatch size $m$, error bound $\delta$, pre-computed correction $C_1(X)$ distribution, initial sample $\theta_1$.}
\Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta_T\}$ from $p(\theta)$\;}
\For{$t=\{1, \ldots, T\}$}{
    Propose a candidate $\theta'$\ from proposal distribution $q(\theta'\mid\theta_t)$\;
    Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(\theta_t,\theta')$ and sample variance $s^2_{\Delta^*})$\;
    Estimate moments $E|\Lambda_i-\Lambda|$ and $E|\Lambda_i-\Lambda|^3$ from the sample, and error $\epsilon$ from Equation~\ref{eq:clt-bounds3}\;
    \While {$s^2_{\Delta^*} \ge 1$ {\bf or} $\epsilon > \delta$}{
    	Draw $m$ more samples to augment the minibatch, update $\Delta^*$, $s^2_{\Delta^*}$ and $\epsilon$ estimates\;
    }
    Draw $X_{\rm nc} \sim \mathcal{N}(0,1-s^2_{\Delta^*})$ and
    $X_{\rm corr}$ from the correction distribution $C_1(X)$\;

        \eIf{$\Delta^* + X_{\rm nc} + X_{\rm corr} > 0$}{
            Accept the candidate, $\theta_{t+1} = \theta'$\;
        }{
            Reject and re-use the old sample, $\theta_{t+1} = \theta_t$\;
        }

@@

The complete algorithm is given in~\ref{alg:our_algorithm}. A few points:
\begin{itemize}
\item The algorithm uses an adaptive step size so as to use the smallest possible
  average minibatch size. Unlike previous work however (and see the experiments section)
  the size distribution is short-tailed.
\item An additional normal variable $X_{\rm nc}$ is added to $\Delta^*$ to produce a variable
  with unit variance. This is not mathematically necessary, but allows us to use a single
  correction distribution $C_1$ for $X_{\rm corr}$, saving on memory footprint.
\end{itemize}

The sample variance $s^2_{\Delta^*}$ is proportional to $||\theta'-\theta||^2$
whose distribution for Normal proposals is the square of a normal variable. 
% {\color{blue} Daniel: TODO improve/describe Table~\ref{tab:xcorr} and try to
% relocate Algorithm~\ref{alg:our_algorithm} to be somewhere nicer.}

@@

For the analysis, in Section~\ref{ssec:delta_star_distribution}, we first
present bounds on the absolute and relative error (in terms of the CDFs) of the
distribution of $\Delta^*$ vs a Gaussian. We then show in
Section~\ref{ssec:preserve_bounds} that these error bounds are preserved after
the addition of other random variables, in particular $X_{\rm nc}$ and $X_{\rm
corr}$. From this it follows that the acceptance test has the same error bound.

@@

%{\color{blue} Daniel: the following lemma is, I think, from Corollary 2
%in~\cite{explicit-clt05} but I cannot derive the exact lemma as we have it
%here.} To start, we use the following quantitative central-limit result
%(from~\cite{explicit-clt05}):

\begin{lemma}\label{lem:quant_clt}
Let $X_1,\ldots,X_n$ be a set of zero-mean, independent, identically-distributed
random variables with sample mean $\bar{X}$ and sample variance $s^2_X$
where:
\begin{equation}
    \bar{X} = \frac{1}{n}\sum_{i=1}^nX_i \qquad {\rm and} \qquad s_X = \frac{1}{n}\left(\sum_{i=1}^n(X_i-\bar{X}^2\right)^{\frac{1}{2}}.
\end{equation}
This implies $t=\bar{X}/s_X$ has an approximate Student's distribution
which approaches a normal distribution in the limit. Then
from~\cite{explicit-clt05}:
\begin{equation}\label{eq:clt-bounds}
    \sup_x|{\rm Pr}(t<x) - \Phi(x)| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}}.
\end{equation}
\end{lemma}

\begin{proof}
  The following bound is given immediately after Corollary 2 in from~\cite{explicit-clt05}:
\begin{equation}\label{eq:clt-boundsx}
-6.4E|X|^3-2E|X| \leq    \sup_x|{\rm Pr}(t<x) - \Phi(x)|{\sqrt{n}} \leq 1.36E|X|^3
\end{equation}
This bound applies to $x\geq 0$. Applying the bound to $-x$ when $x<0$
and combining with $x>0$, we obtain the weaker but unqualified bound
in equation (\ref{eq:clt-bounds}).
\end{proof}

%% {\color{blue} Daniel: the notation here was (and still is, to a lesser extent)
%% the source of considerable confusion for me. Do we want $\sigma(X)$ or
%% $\sigma(\hat{X})$? I think in Lemma~\ref{lem:quant_clt} we want the former, not
%% the latter. OH ALSO, this section uses $n$ as the full data size, which means
%% before this section, we should use $n$, not $N$, for the full data size, right?
%% The Bardenet paper uses $n$ to represent the full data size.

%% JFC - this bound applies to minibatches, not the full dataset so the size $n$ is
%% deliberately not $N$.

Lemma~\ref{lem:quant_clt} demonstrates that as long as we know the first and
third absolute moments $E|X|$ and $E|X|^3$, we can bound the error of the normal
approximation, which decays as $O(n^{-\frac{1}{2}})$. Making the change of
variables $y = x s_X)$, Equation~\ref{eq:clt-bounds} becomes
\begin{equation}\label{eq:clt-bounds2}
   \sup_y\left|{\rm Pr}(\bar{X}< y) - \Phi\left(\frac{y}{s_X}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}},
\end{equation}
showing that the distribution of $\bar{X}$ approaches the normal distribution
$\mathcal{N}(0,s_X)$ whose variance is $s_X$,
measured from the sample.

To apply this to our test, let $X_i = \Lambda_i(\theta,\theta') -
\Lambda(\theta,\theta')$, so that the $X_i$ are zero-mean, i.i.d. variables. If
instead of all $n$ samples, we only extract a subset of $b$ samples
corresponding to our minibatch, we can connect $\bar{X}$ with our $\Delta^*$
term:
\begin{equation}\label{eq:x-hat}
    \bar{X} = \Delta^*(\theta,\theta') - \Delta(\theta,\theta'),
\end{equation}
so that $s_X = s_{\Delta^*}$.
%{\color{blue} Daniel: This time it
%  makes sense to use $\sigma(\hat{X})$ instead of $\sigma(X)$.}
This results in the following Corollary:

\begin{corollary}\label{cor:our_bound_delta_prime}
We can now substitute into Equation~\ref{eq:clt-bounds2} and displace by the mean, giving:
\begin{equation}\label{eq:clt-bounds3}
    \sup_y\left|{\rm Pr}(\Delta^*< y) - \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{b}} = \epsilon(\theta,\theta',b)
\end{equation}
\end{corollary}

Corollary~\ref{cor:our_bound_delta_prime} shows that the distribution of
$\Delta^*$ approximates a Normal distribution with mean $\Delta$ and variance
$s^2_{\Delta^*}$. Furthermore, it bounds the error with \emph{estimable
quantities}: both $E|X|$ and $E|X|^3$ can be estimated as means of $|\Lambda_i
- \Lambda|$ and $|\Lambda_i - \Lambda|^3$, respectively, on each minibatch. We
expect this will often be accurate enough on minibatches with hundreds or
thousands of points, but otherwise bootstrap CIs can be computed from those
sequences. Since the bounds are monotone in $E|X|$ and $E|X|^3$, using upper
bootstrap CI limits will provide high-confidence error bounds.

@@

%% {\color{blue} Daniel: Hold on, the following Corollary seems to add $X_{\rm norm}$
%% to both random variables we are comparing (and \emph{then} $X_{\rm corr}$). But
%% didn't the previous section already justify that $\Delta^* \approx \Delta +
%% X_{\rm norm}$? That is, I don't think we need to add another $X_{\rm norm}$; I
%% think we just want to add $X_{\rm corr}$. Otherwise, this new test does not
%% match what we wrote earlier.

%% I think the easiest way to resolve this is to do the following: we assume that
%% $\Delta^*$ is already $\Delta + X_{\rm noise}$. Then all we need to do is add
%% $X_{\rm corr}$. So this is the same as what we are doing, except we do not need
%% to apply Lemma~\ref{lem:cdf_bounds} twice; we just need to do it once because we
%% are only adding $X_{\rm corr}$, not $X_{\rm norm}$.}

@@

\begin{corollary}\label{cor:bounds_preserved}
If $\sup_y|{\rm Pr}(\Delta^* < y) - \Phi(\frac{y-\Delta}{s_{\Delta^*}})|
\leq \epsilon(\theta,\theta',b)$, then
\begin{equation}
    \sup_y|{\rm Pr}(\Delta^*+X_{\rm nc}+X_{\rm corr} < y) - S(y-\Delta)| \leq \epsilon(\theta,\theta',b)
\end{equation}
where $S(x)$ is the standard logistic function, and $X_{\rm nc}$ and $X_{\rm corr}$ are generated as per Algorithm 1. 
\end{corollary}

\begin{proof}
We apply Lemma~\ref{lem:cdf_bounds} twice. First take:
\begin{equation}
    P(y) = {\rm Pr}(\Delta^* < y) \qquad {\rm and}\qquad Q(y) = \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)
\end{equation}
and convolve with the distribution of $X_n$ which has density
%% {\color{blue}
%%   (Daniel: to be clear I
%% think this was supposed to be $X_{\rm norm}$ instead of $X_n$ but see comments
%% earlier about why I think it is not necessary.)}
$\phi(X/\sigma_n)$ where $\sigma_n^2 = 1 - s^2_{\Delta^*}$. This yields the
next iteration of $P$ and $Q$:
\begin{equation}
    P'(y) = {\rm Pr}(\Delta^*+X_{\rm nc} < y) \qquad {\rm and}\qquad Q'(y) = \Phi\left({y-\Delta}\right)
\end{equation}
Now we convolve with the distribution of $X_{\rm corr}$ which gives:
\begin{equation}
    P''(y) = {\rm Pr}(\Delta^*+X_{\rm nc}+X_{\rm corr} < y) \qquad {\rm and}\qquad Q''(y) = S\left({y-\Delta}\right)
\end{equation}
Both steps preserve the error bound $\epsilon(\theta,\theta',b)$. Finally
$S(y-\Delta)$ is a logistic CDF centered at $\Delta$, and so $S(y-\Delta) = {\rm
Pr}(\Delta + X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We conclude
that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm
nc}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm
log} > 0)$ by at most $\epsilon$.
\end{proof}

@@

%% {\color{blue} Daniel: Do we want to include this new analysis? There seem to be
%% two different ways we show the bound. The first part (above) is nice and I like
%% it (and it seems to show all we need for Algorithm 1, doesn't it?}

@@

Here the limit $G_n(x,y)$ is a normal CDF plus a correction term that decays as
$n^{1/2}$.
%{\color{blue} (Daniel: how did the equation in the next sentence
% appear? I verified it and it seems correct, but it looks like it popped out of
%nowhere.)}
% JFC: It comes from taking a second derivative. You dont need to explain it. 
Importantly, since $\phi^{''}(x) = x^2\phi(x) - \phi(x)$ where
$\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
\begin{equation}\label{eq:GNderivatives}
    G_n(x,y) = \Phi(x) + \frac{y}{6\sqrt{n}}(2\phi^{''}(x)+3\phi(x))
\end{equation}
From which we see that $G_n(x,y)$ is a linear combination of $\Phi(x)$,
$\phi(x)$ and $\phi^{''}(x)$. This is quite fortuitous. In Algorithm 1, we
correct for the difference in $\sigma$ between $\Delta^*$ and the variance
needed by $X_{\rm corr}$ using $X_{\rm nc}$. This same method works when we
wish to estimate the error in $\Delta^*$ vs $G_n(x,y)$. Since all of the
component functions of $G_n(x,y)$ are derivatives of a (unit variance)
$\Phi(x)$, adding a normal variable with variance $\sigma'$ increases the
variance of all three functions to $1+\sigma'$. Thus we add $X_{\rm nc}$ as
per Algorithm 1 preserving the limit in Equation~\ref{eq:GNderivatives}.
